---
title: Classwork 3
subtitle: Property Rights, the Coase Theorem, and the Environment
date: 2025-09-19
code-fold: true
execute: 
  eval: true
  echo: false
  warning: false
  message: false
from: markdown+emoji
---
```{r}
#| include: false

library(knitr)
library(rmarkdown)
library(tidyverse)
library(skimr)
library(ggthemes)
library(hrbrthemes)

theme_set(theme_ipsum() +
          theme(strip.background =element_rect(fill="lightgray"),
                axis.title.x = element_text(size = rel(1.5)),
                axis.title.y = element_text(size = rel(1.5)),
                legend.title = element_text(size=rel(1.25))
                ))
```



# Question 1. Property Rights and Coasian Bargaining

Consider a society with just two people. *Person 1* enjoys playing loud music for $h$ hours, while *Person 2* dislikes the noise. Both are **self-interested** and care only about their own payoff.

- *Person 1*’s marginal benefit (in dollars) from music is:  
  $MB(h) = 4 - 2h$.  
- *Person 2*’s marginal cost (in dollars) from music is:  
  $MC(h) = 2h$.  

:::callout-note
Here, $MB(h)$ and $MC(h)$ are functions of $h$.  
Transfers $T$ (in dollars) are lump-sum payments between the two people.  
:::


## a. Market outcome
If the decision is left entirely to *Person 1*, what level of music $h_M$ will she choose? Why?

**Example Answer**  
```{r}
MB <- function(h) 4 - 2*h
MC <- function(h) 2*h

df <- data.frame(h = seq(0, 2.2, by = 0.01))
df$MB <- MB(df$h)
df$MC <- MC(df$h)

h_star <- 1
h_M    <- 2

ggplot(df, aes(h)) +
  geom_hline(yintercept = 0, linewidth = 1.1) +
  geom_vline(xintercept = 0, linewidth = 1.1) +
  geom_line(aes(y = MB), linewidth = 1.1, color = 'blue') +
  geom_vline(xintercept = h_M,    linetype = "dashed") +
  annotate("text", x = h_M,    y = 3.75, label = "h_M = 2", vjust = -0.5, angle = 90) +
  labs(title = "Marginal Benefit of Loud Music",
       x = "h (hours of loud music)", y = "Dollars") +
  theme_ipsum(base_size = 13)
```


Person 1 chooses $h$ to maximize her own payoff and ignores Person 2’s cost.  
She sets $MB(h)=0$:  
$$
4-2h=0 \;\;\Rightarrow\;\; \boxed{h_M=2}.
$$  


```{r}
MB <- function(h) 4 - 2*h
MC <- function(h) 2*h

df <- data.frame(h = seq(0, 2.2, by = 0.01))
df$MB <- MB(df$h)
df$MC <- MC(df$h)

h_star <- 1
h_M    <- 2

ggplot(df, aes(h)) +
  geom_hline(yintercept = 0, linewidth = 1.1) +
  geom_vline(xintercept = 0, linewidth = 1.1) +
  geom_line(aes(y = MB), linewidth = 1.1, color = 'blue') +
  geom_vline(xintercept = h_M,    linetype = "dashed") +
  geom_polygon(
    data = data.frame(
      h = c(0, 2, 0),
      y = c(0, 0, 4)
    ),
    aes(x = h, y = y),
    fill = "blue", alpha = 0.3
  ) +
  annotate("text", x = h_M,    y = 3.75, label = "h_M = 2", vjust = -0.5, angle = 90) +
  labs(title = "Total Benefit of Loud Music",
       x = "h (hours of loud music)", y = "Dollars") +
  theme_ipsum(base_size = 13)
```

Total benefit ($TB$) is the area under the $MB$ curve. So, total benefit is maximized at $h=2$.

**Why?** With no private cost, she keeps increasing $h$ until her marginal benefit hits zero; the neighbor’s harm is not internalized.

*Examples*

- At $h=0$: Total benefit of playing music is zero.

- At $h=1$: Total benefit is smaller than the above:
```{r}
MB <- function(h) 4 - 2*h
MC <- function(h) 2*h

df <- data.frame(h = seq(0, 2.2, by = 0.01))
df$MB <- MB(df$h)
df$MC <- MC(df$h)

h_star <- 1
h_M    <- 2

ggplot(df, aes(h)) +
  geom_hline(yintercept = 0, linewidth = 1.1) +
  geom_vline(xintercept = 0, linewidth = 1.1) +
  geom_line(aes(y = MB), linewidth = 1.1, color = 'blue') +
  geom_vline(xintercept = h_star,    linetype = "dashed") +
  geom_polygon(
    data = data.frame(
      h = c(0, 1, 1, 0),
      y = c(0, 0, 2, 4)
    ),
    aes(x = h, y = y),
    fill = "blue", alpha = 0.3
  ) +
  annotate("text", x = h_star,    y = 3.75, label = "h = 1", vjust = -0.5, angle = 90) +
  labs(title = "Total Benefit of Loud Music when h = 1",
       x = "h (hours of loud music)", y = "Dollars") +
  theme_ipsum(base_size = 13)
```


<br><br>


## b. Socially efficient outcome
What is the socially optimal level $h^{*}$? Why?


**Example Answer**  


```{r}
MB <- function(h) 4 - 2*h
MC <- function(h) 2*h

df <- data.frame(h = seq(0, 2.2, by = 0.01))
df$MB <- MB(df$h)
df$MC <- MC(df$h)

h_star <- 1
h_M    <- 2

ggplot(df, aes(h)) +
  geom_hline(yintercept = 0, linewidth = 1.1) +
  geom_vline(xintercept = 0, linewidth = 1.1) +
  geom_line(aes(y = MB), linewidth = 1.1, color = 'blue') +
  geom_line(aes(y = MC), linewidth = 1.1, color = 'red') +
  geom_point(data = data.frame(h = 1, y = 2),
             aes(x=h, y=y), size = rel(2)) +
  geom_vline(xintercept = h_star, linetype = "dashed") +
  annotate("text", x = h_star, y = 3.9, label = "h* = 1", vjust = -0.5, angle = 90) +
  labs(title = "Marginal Benefit and Marginal Cost of Loud Music",
       x = "h (hours of loud music)", y = "Dollars") +
  theme_ipsum(base_size = 13)
```


Efficiency requires $MB(h)=MC(h)$, so that social welfare (total surplus) is maximized: 

$$
4-2h=2h \Rightarrow \boxed{h^*=1}
$$.


**Why?** At $h=1$, the next unit’s benefit equals the next unit’s cost, so total surplus is maximized:  
$$
TS(h)= TB(h) - TC(h),
$$  
which peaks at $h=1$.


The total surplus of loud music is the triangle area below:
```{r}
MB <- function(h) 4 - 2*h
MC <- function(h) 2*h

df <- data.frame(h = seq(0, 2.2, by = 0.01))
df$MB <- MB(df$h)
df$MC <- MC(df$h)

h_star <- 1
h_M    <- 2

ggplot(df, aes(h)) +
  geom_hline(yintercept = 0, linewidth = 1.1) +
  geom_vline(xintercept = 0, linewidth = 1.1) +
  geom_line(aes(y = MB), linewidth = 1.1, color = 'blue') +
  geom_line(aes(y = MC), linewidth = 1.1, color = 'red') +
  geom_point(data = data.frame(h = 1, y = 2),
             aes(x=h, y=y), size = rel(2)) +
  geom_vline(xintercept = h_star, linetype = "dashed") +
  geom_polygon(
    data = data.frame(
      h = c(0, 1, 0),
      y = c(0, 2, 4)
    ),
    aes(x = h, y = y),
    fill = "green", alpha = 0.3
  ) +
  annotate("text", x = h_star, y = 3.9, label = "h* = 1", vjust = -0.5, angle = 90) +
  labs(title = "Total Surplus of Loud Music",
       x = "h (hours of loud music)", y = "Dollars") +
  theme_ipsum(base_size = 13)
```

This is because the total benefit is the area under the MB, and the total cost is the area under the MC:

- Total benefit ($TB$) from 1 hour of loud music
```{r}
MB <- function(h) 4 - 2*h
MC <- function(h) 2*h

df <- data.frame(h = seq(0, 2.2, by = 0.01))
df$MB <- MB(df$h)
df$MC <- MC(df$h)

h_star <- 1
h_M    <- 2

ggplot(df, aes(h)) +
  geom_hline(yintercept = 0, linewidth = 1.1) +
  geom_vline(xintercept = 0, linewidth = 1.1) +
  geom_line(aes(y = MB), linewidth = 1.1, color = 'blue') +
  geom_line(aes(y = MC), linewidth = 1.1, color = 'red') +
  geom_point(data = data.frame(h = 1, y = 2),
             aes(x=h, y=y), size = rel(2)) +
  geom_vline(xintercept = h_star, linetype = "dashed") +
  geom_polygon(
    data = data.frame(
      h = c(0, 1, 1, 0),
      y = c(0, 0, 2, 4)
    ),
    aes(x = h, y = y),
    fill = "blue", alpha = 0.3
  ) +
  annotate("text", x = h_star, y = 3.9, label = "h* = 1", vjust = -0.5, angle = 90) +
  labs(title = "Total Benefit of Loud Music",
       x = "h (hours of loud music)", y = "Dollars") +
  theme_ipsum(base_size = 13)
```

- Total cost ($TC$) from 1 hour of loud music
```{r}
MB <- function(h) 4 - 2*h
MC <- function(h) 2*h

df <- data.frame(h = seq(0, 2.2, by = 0.01))
df$MB <- MB(df$h)
df$MC <- MC(df$h)

h_star <- 1
h_M    <- 2

ggplot(df, aes(h)) +
  geom_hline(yintercept = 0, linewidth = 1.1) +
  geom_vline(xintercept = 0, linewidth = 1.1) +
  geom_line(aes(y = MB), linewidth = 1.1, color = 'blue') +
  geom_line(aes(y = MC), linewidth = 1.1, color = 'red') +
  geom_point(data = data.frame(h = 1, y = 2),
             aes(x=h, y=y), size = rel(2)) +
  geom_vline(xintercept = h_star, linetype = "dashed") +
  geom_polygon(
    data = data.frame(
      h = c(0, 1, 1),
      y = c(0, 0, 2)
    ),
    aes(x = h, y = y),
    fill = "red", alpha = 0.3
  ) +
  annotate("text", x = h_star, y = 3.9, label = "h* = 1", vjust = -0.5, angle = 90) +
  labs(title = "Total Cost of Loud Music",
       x = "h (hours of loud music)", y = "Dollars") +
  theme_ipsum(base_size = 13)
```





## c. Person 1 has the right to play
Suppose the government gives *Person 1* the legal right to play as much as she wants ($h=h_M$).  
*Person 2* may then “bribe” *Person 1* to reduce music in exchange for a lump-sum transfer $T$ (in dollars).  
Assume there are no transaction costs in bargaining.

- Show how bargaining can still lead them to agree on the socially optimal level $h^{*}$.  
  - Identify the possible range of transfers $T$ (in dollars), paid from *Person 2* to *Person 1*, that makes both better off compared to their initial legal entitlement ($h=h_M, T=0$).

**Example Answer**  
Starting from $h=2$, 
$$
MC(2)=4>MB(2)=0,
$$

so both gain by reducing $h$ until $MB=MC$, i.e., to $\boxed{h^{*}=1}$.

Transfers to cut noise from $2\to 1$: 

- Person 2 should be better off from the bargaining.
  - This means that Person 2’s total cost (harm from music plus any transfer paid to Person 1) at $h=1$ with transfer $T$ must be no greater than at the initial entitlement $h=2$ with $T=0$.

$$
\begin{align}
TC(h=1) + T &\;\leq\; TC(h=2) \qquad \Leftrightarrow\\
\left(\frac{1}{2}\times 1 \times 2\right) + T &\;\leq\; \frac{1}{2}\times 2 \times 4 \qquad \Leftrightarrow\\
1 + T &\;\leq\; 4\qquad \Leftrightarrow\\
T&\;\leq\; 3
\end{align}
$$

- Person 1 should also be better off from the bargaining. 
  - This means that Person 1’s net payoff (the benefit from music plus any transfer received) at $h=1$ with $T$ must be at least as high as at the initial entitlement $h=2$ with $T=0$.
  
$$
\begin{align}
TB(h=1) + T &\;\geq\; TB(h=2) \qquad \Leftrightarrow\\
\left(\frac{1}{2}\times (4 + 2) \times 1\right) + T &\;\geq\; \frac{1}{2}\times 2 \times 4 \qquad \Leftrightarrow\\
3 + T &\;\geq\; 4\qquad \Leftrightarrow\\
T&\;\geq\; 1
\end{align}
$$
Therefore, any transfer $T$ with
$$
1\;\leq\; T \;\leq\; 3
$$
paid from Person 2 to Person 1 makes both (weakly) better off.  


<br><br>

## d. Person 2 has the right to silence
Now suppose the government gives *Person 2* the legal right to silence ($h=0$).  
*Person 1* may then “bribe” *Person 2* to allow some music in exchange for a lump-sum transfer $T$ (in dollars).  
Assume there are no transaction costs in bargaining.

- Show how bargaining can still lead them to agree on the socially optimal level $h^{*}$.  
  - Identify the possible range of transfers $T$ (in dollars), paid from *Person 1* to *Person 2*, that makes both better off compared to their initial legal entitlement ($h=0, T=0$).


**Example Answer**  
Starting from $h=0$, 
$$
MC(0)=0<MB(0)=4,
$$

so both gain by increasing $h$ until $MB=MC$, i.e., to $\boxed{h^{*}=1}$.

Transfers to increase noise from $0\to 1$: 

- Person 2 should be better off from the bargaining.
  - This means that Person 2’s net payoff (transfer received from Person 1 minus the harm from music) at $h=1$ with transfer $T$ must be at least as great as under the initial entitlement $h=0$ with $T=0$.

$$
\begin{align}
T - TC(h=1)  &\;\geq\; -TC(h=0) \qquad \Leftrightarrow\\
T - \left(\frac{1}{2}\times 1 \times 2\right)  &\;\geq\; 0 \qquad \Leftrightarrow\\
T   &\;\geq\; 1
\end{align}
$$

- Person 1 should also be better off from the bargaining. 
  - This means that Person 1’s net payoff (the benefit from music minus any transfer paid to person 2) at $h=1$ with $T$ must be at least as high as at the initial entitlement $h=0$ with $T=0$.
  
$$
\begin{align}
TB(h=1) - T &\;\geq\; TB(h=0) \qquad \Leftrightarrow\\
\left(\frac{1}{2}\times (4 + 2) \times 1\right) - T &\;\geq\; 0 \qquad \Leftrightarrow\\
3  &\;\geq\; T
\end{align}
$$
Therefore, any transfer $T$ with
$$
1\;\leq\; T \;\leq\; 3
$$
paid from Person 1 to Person 2 makes both (weakly) better off.  


<br><br>

## Why transfers don’t justify stopping at $h = 0.5$ or $h = 1.5$?

- The **total pie** from bargaining is the **total surplus**—the total dollar value created, equal to Person 1’s total benefit from music minus Person 2’s total harm.  
  - Intuitively, the **total surplus** represents the size of the pie before deciding how to slice it between the two people.  
  - We call it a “pie” because it measures the overall gains available from negotiation: the larger the pie, the more there is for the two to share, even though bargaining determines who gets which slice.  
  
$$
TS(h) = TB(h) - TC(h)
$$

- Lump-sum transfers ($T$) only redistribute this surplus between the two people; they don’t create or destroy it.  

- If Person 1 has the right to play:  
$$
\begin{align}
(\text{Total Pie}) &= (TB(h) + T) - (TC(h) + T)\\
&= TS(h)
\end{align}
$$

- If Person 2 has the right to silence:  
$$
\begin{align}
(\text{Total Pie}) &= (TB(h) - T) - (TC(h) + T)\\
&= TS(h)
\end{align}
$$

-	At $h=0.5$, there’s unclaimed surplus left on the table. Moving closer to $h=1$ increases total pie size before dividing it.
-	At $h=1.5$, they’re wasting surplus (total harm > total benefit). Moving closer to $h=1$ makes the pie bigger for both.

So regardless of how the pie is sliced (what $T$ is), the pie is maximized only at $h^{*}=1$.

How the pie is divided—i.e., the actual transfer size—depends on their relative bargaining power.