---
title: Classwork 13
subtitle: Self-test for Midterm Exam II
date: 2025-11-19
code-fold: true
execute:
  echo: false
  eval: true
  warning: false
  message: false
from: markdown+emoji
---
```{r}
#| include: false

library(knitr)
library(rmarkdown)
library(tidyverse)
library(skimr)
library(ggthemes)
library(hrbrthemes)

theme_set(theme_ipsum() +
          theme(strip.background =element_rect(fill="lightgray"),
                axis.title.x = element_text(size = rel(1.5)),
                axis.title.y = element_text(size = rel(1.5)),
                legend.title = element_text(size=rel(1.25))
                ))
```

# Section 1. Multiple Choice

## Question 1
What is the core mechanism behind bounded willpower?

a. Social comparison
b. Loss aversion
c. Present bias causing time-inconsistent preferences
d. Failure to understand operating-cost differences

<br>

## Question 2
A $250 benefit arrives 20 years in the future. All else equal, which change will increase its present value?

a. Increasing the discount rate from 3% to 5%
b. Shortening the time horizon from 20 years to 10 years
c. Moving from real to nominal dollars
d. Converting the benefit into a cost

<br>

## Question 3
Which statement is most accurate?

a. A policy with positive $PV_{NB}$ must be both efficient and equitable
b. A policy with negative $PV_{NB}$ can still be equitable
c. Efficiency always implies equity
d. Equity is automatically satisfied when the B/C ratio exceeds 1


<br>

## Question 4
In a cost-effective pollution reduction policy with multiple firms:

a. All firms must abate the same quantity.
b. Each firm must face the same average abatement cost.
c. Total abatement is zero.
d. The marginal abatement costs are equalized across firms.


<br>

## Question 5
Which of the following most directly describes carbon leakage?

a. Emissions reductions that occur due to innovation spillovers.
b. Emissions being shifted abroad when production relocates to countries with weaker climate policy.
c. Emissions reductions concentrated in non-traded sectors.
d. The leakage of methane from oil and gas wells.


<br>



# Section 2. Filling-in-the-Blanks

## Question 6
Effective policy design requires integrating four components: Incentives, ________, Institutions, and ________.


<br>


## Question 7
The IPCC combines Climate Models (physical simulations) with Socioeconomic Scenarios to form ________________________, which represent different "what if" futures ranging from sustainability to fossil-fueled development.

<br>


## Question 8
The Paris Agreement includes a mechanism called the ________________________, occurring every 5 years, which evaluates collective progress to inform the next round of Nationally Determined Contributions (NDCs).

<br>


# Section 4. Analytical Modeling & Critique

## Question 9

Consider a perfectly competitive market for **solar energy** that generates **positive externalities** such as reduced carbon emissions and knowledge spillovers.  
The **private marginal benefit (PMB)** and **marginal cost (MC)** of solar production are:

$$
\begin{aligned}
PMB(Q) &= 110 - 2Q\\
MC(Q) &= 20 + Q
\end{aligned}
$$

and the **external marginal benefit (EMB)** is:

$$
EMB(Q) = 60
$$



### Part A
Find the **market equilibrium** quantity $Q_M$ and price $P_M$.  
Show all steps.

<br>

### Part B
Compute the **social marginal benefit (SMB)** function.

<br>


### Part C
Find the **socially optimal output** $Q^*$ and price $P^*$.

<br>


### Part D
- Graph the $PMB$, $SMB$, and $MC$ curves.  
  - Label both the **market equilibrium** ($Q_M, P_M$) and the **socially optimal equilibrium** ($Q^*, P^*$).

- Graphically shade and label the **deadweight loss (DWL)**  caused by the positive externality on the graph.


```{r}
#| fig-width: 7
#| fig-height: 6
#| fig-align: center
#| echo: false

x <- seq(0, 90, 10)
y <- seq(0, 180, 20)
ggplot(data = data.frame(x =x,
                         y = x)) +
  geom_point(aes(x=x,y=y), size = 0, color = 'white') +
  geom_vline(xintercept = 5, color = 'grey80') +
  geom_vline(xintercept = 10, color = 'grey80') +
  geom_vline(xintercept = 15, color = 'grey80') +
  geom_vline(xintercept = 20, color = 'grey80') +
  geom_vline(xintercept = 25, color = 'grey80') +
  geom_vline(xintercept = 30, color = 'grey80') +
  geom_vline(xintercept = 35, color = 'grey80') +
  geom_vline(xintercept = 40, color = 'grey80') +
  geom_vline(xintercept = 45, color = 'grey80') +
  geom_vline(xintercept = 50, color = 'grey80') +
  geom_vline(xintercept = 55, color = 'grey80') +
  geom_vline(xintercept = 60, color = 'grey80') +
  geom_vline(xintercept = 65, color = 'grey80') +
  geom_vline(xintercept = 70, color = 'grey80') +
  geom_vline(xintercept = 75, color = 'grey80') +
  geom_vline(xintercept = 80, color = 'grey80') +
  geom_vline(xintercept = 85, color = 'grey80') +
  geom_vline(xintercept = 90, color = 'grey80') +
  geom_hline(yintercept = 10, color = 'grey80') +
  geom_hline(yintercept = 20, color = 'grey80') +
  geom_hline(yintercept = 30, color = 'grey80') +
  geom_hline(yintercept = 40, color = 'grey80') +
  geom_hline(yintercept = 50, color = 'grey80') +
  geom_hline(yintercept = 60, color = 'grey80') +
  geom_hline(yintercept = 70, color = 'grey80') +
  geom_hline(yintercept = 80, color = 'grey80') +
  geom_hline(yintercept = 90, color = 'grey80') +
  geom_hline(yintercept = 100, color = 'grey80') +
  geom_hline(yintercept = 110, color = 'grey80') +
  geom_hline(yintercept = 120, color = 'grey80') +
  geom_hline(yintercept = 130, color = 'grey80') +
  geom_hline(yintercept = 140, color = 'grey80') +
  geom_hline(yintercept = 150, color = 'grey80') +
  geom_hline(yintercept = 160, color = 'grey80') +
  geom_hline(yintercept = 170, color = 'grey80') +
  geom_hline(yintercept = 180, color = 'grey80') +
  geom_hline(yintercept = 0) +
  geom_vline(xintercept = 0) +
  scale_x_continuous(breaks = x,
                     labels = x) +
  scale_y_continuous(breaks = y,
                     labels = y) +
  labs(x = 'Q', y= 'P') 
```



<br>

### Part E

Calculate the **optimal per-unit subsidy** that would achieve the socially efficient outcome.

<br>



### Part F

Compute the **market price with the subsidy**, i.e., the prices received by producers and paid by consumers.



<br>



### Part G
Compute the **subsidy expenditure** and the **change in total welfare** from implementing the policy.

- Subsidy expenditure represents the **total fiscal cost** to the government of providing the per-unit subsidy.  



<br>
