---
title: Classwork 4
subtitle: Optimal Harvesting in a Common-pool Fishery
date: 2025-09-26
code-fold: true
execute: 
  echo: false
  eval: true
  warning: false
  message: false
from: markdown+emoji
---
```{r}
#| include: false

library(knitr)
library(rmarkdown)
library(tidyverse)
library(skimr)
library(ggthemes)
library(hrbrthemes)

theme_set(theme_ipsum() +
          theme(strip.background =element_rect(fill="lightgray"),
                axis.title.x = element_text(size = rel(1.5)),
                axis.title.y = element_text(size = rel(1.5)),
                legend.title = element_text(size=rel(1.25))
                ))
```


<br>

Consider a fishery where the total benefit (**TB**) and total cost (**TC**) of fishing effort $E$ are:

- $TB(E)=1.5E - E^{2}$ (in \$)
- $TC(E)=0.5E$ (in \$)

where $E$ is the level of fishing effort (e.g., number of boat-days).

## Key Concepts

Before beginning, recall these definitions:

- **Average Benefit (AB):**  
  $AB(E) = \dfrac{TB(E)}{E}$.  
  In economics, AB measures benefit per unit of input; here, it shows the **revenue per boat-day**.  
  *For this fishery:* $AB(E) = \dfrac{1.5E - E^2}{E} = 1.5 - E$.  

- **Average Cost (AC):**  
  $AC(E) = \dfrac{TC(E)}{E}$.  
  In economics, AC measures cost per unit of input; here, it is the **cost per boat-day**.  
  *For this fishery:* $AC(E) = \dfrac{0.5E}{E} = 0.5$.  

- **Marginal Benefit (MB):**  
  $MB(E) = \dfrac{dTB}{dE}$.  
  In economics, MB is the extra benefit from one more unit of input; here, the **additional revenue from one more boat-day**.  
  *For this fishery:* $MB(E) = 1.5 - 2E$.  

- **Marginal Cost (MC):**  
  $MC(E) = \dfrac{dTC}{dE}$.  
  In economics, MC is the extra cost from one more unit of input; here, the **additional cost of adding one boat-day**.  
  *For this fishery:* $MC(E) = 0.5$.  
  
<br>

## Part A — Total Curves
Draw the **Total Benefit (TB)** and **Total Cost (TC)** curves as functions of fishing effort $E$.  Place these curves in the **top panel** of a two-panel figure.

## Part B — Marginal Curves
On the **bottom panel**, draw the **Marginal Benefit (MB)** and **Marginal Cost (MC)** curves as functions of fishing effort $E$.  

## Part C — Efficient and Open-Access Effort Levels
Make sure the top and bottom panels share the same horizontal scale for effort $E$, so the relationship between total and marginal spaces is consistent.

- Recall that:
  - $MB(E)=\dfrac{d\,TB}{dE}=1.5 - 2E$
  - $MC(E)=\dfrac{d\,TC}{dE}=0.5$

On both panels, **mark and label** the two benchmark effort levels using **vertical guide lines** that align across panels:

- The **efficient effort** $E^*$:
  Draw a vertical line at $E^*$ on **both** the top (TB/TC) and bottom (AB/AC/MB/MC) panels and label it “$E^*$”.

- The **open-access effort** $E_{OA}$:
  Draw a vertical line at $E_{OA}$ on **both** panels and label it “$E_{OA}$”.


## Part D — Intuition Behind $E^*$ and $E_{OA}$
Explain the intuition behind these two effort levels:

1. **Why does $E^*$ arise?**  
2. **Why does $E_{OA}$ arise under open access?**  


## Part E — Policy Intervention with a License Fee
- **License fee:** Explain, both analytically and intuitively, how charging a **per-unit license fee** (e.g., per boat-day) can reduce fishing effort from the open-access level $E_{OA}$ down to the efficient level $E^*$.  
- Explain why it is important to align individual fishers’ incentives with the socially optimal level of effort in a fishery. In your answer, discuss:
	1.	**Incentive problem**: Why do individual fishers’ private decisions differ from what is best for the group?
	2.	**Economic rent**: What happens to economic rents under open access when incentives are not aligned?
	3.	**Sustainability**: Why does aligning incentives matter for the long-run health of both the fish stock and the fishing economy?

```{r}
#| eval: false
# Load packages
library(tidyverse)

# Domain for effort (adjust as needed)
E_max <- 1.5
E <- seq(0, E_max, length.out = 501)

# Define functions
TB <- 1.5*E - E^2
TC <- 0.5*E
MB <- 1.5 - 2*E
MC_AC <- rep(0.5, length(E))
AB <- 1.5 - E


# Data frames
df_level <- tibble(E, TB, TC) |> pivot_longer(cols = c(TB, TC), names_to = "curve", values_to = "value")
df_marg  <- tibble(E, AB, MB, MC_AC) |> pivot_longer(cols = c(AB, MB, MC_AC), names_to = "curve", values_to = "value") |> mutate(curve = ifelse(curve == 'MC_AC', 'AC & MC', curve))

# (a) TB vs TC
ggplot(df_level, aes(x = E, y = value, linetype = curve)) +
  geom_line(size = 1) +
  labs(title = "Total Benefit and Total Cost vs Effort",
       x = "Effort (E)", y = "Dollars", linetype = "Curve") +
  theme_minimal(base_size = 12)

# (b) MB vs MC
ggplot(df_marg, aes(x = E, y = value, linetype = curve)) +
  geom_line(size = 1) +
  labs(title = "Average/Marginal Benefit and Average/Marginal Cost vs Effort",
       x = "Effort (E)", y = "Dollars per unit of E", linetype = "Curve") +
  theme_minimal(base_size = 12)
```


