---
title: Classwork 7
subtitle: Quiz 1
date: 2025-10-08
code-fold: true
execute: 
  echo: false
  eval: true
  warning: false
  message: false
from: markdown+emoji
---
```{r}
#| include: false

library(knitr)
library(rmarkdown)
library(tidyverse)
library(skimr)
library(ggthemes)
library(hrbrthemes)

theme_set(theme_ipsum() +
          theme(strip.background =element_rect(fill="lightgray"),
                axis.title.x = element_text(size = rel(1.5)),
                axis.title.y = element_text(size = rel(1.5)),
                legend.title = element_text(size=rel(1.25))
                ))
```


# Section 1. Multiple Choice and Filling-in-the-Blanks

## Question 1. 
The Coase Theorem states that if property rights are well defined and transaction costs are negligible, then private bargaining will lead to: 

a. A more equitable distribution of wealth.
b. The lowest possible transaction costs.
c. Efficient allocation of resources, even with externalities.
d. Government intervention to correct market failures.



::: {.callout-tip collapse="true"}
### Show answer
**c.** Efficient allocation of resources, even with externalities.

**Why?** With zero transaction costs and well-defined rights, parties can bargain to internalize externalities and reach the efficient outcome regardless of the initial allocation of rights.

:::

## Question 2.
For a (pure) public good, individual willingness to pay (WTP) typically ____(a)____ the true Marginal Benefit (MB), resulting in ____(b)____ when provided through the market.

a. (a) understates; (b) underprovision
b. (a) overstates; (b) overprovision
c. (a) understates; (b) overprovision
d. (a) overstates; (b) underprovision


::: {.callout-tip collapse="true"}
### Show answer
**a.** (a) *understates*; (b) *underprovision*.

**Why?** People have an incentive to free‑ride, so their stated WTP is below their true MB. Summing private demand understates the social MB, yielding too little provision in markets.

:::


## Question 3. 
The two primary characteristics used to classify goods and resources in the table are _________________________ (Does one user's harvest reduce what's left for others?) and _________________________ (Can users who haven't paid be kept out?). 


::: {.callout-tip collapse="true"}
### Show answer

**Rivalry** and **Excludability**.

:::



<!-- ## Question 4. -->
<!-- Elinor Ostrom's work challenged the notion of the "tragedy of the commons" by showing that many communities succeed through _________________________ without needing full privatization or heavy state control. -->

<!-- $\vspace{3cm}$ -->


# Section 2. Analytical Models

Consider a perfectly competitive market for a good whose production generates pollution as a byproduct. The marginal cost ($MC$) of producing the good is given by 

$$
MC(Q) = 10 + 2Q,
$$ 

where $Q$ is the quantity produced. The external marginal cost ($EMC$) of pollution is given by 

$$
EMC(Q) = 2Q.
$$ 

The market demand is 

$$
MB(Q) = 70 - 2Q.
$$



## Question 4
Draw the $MB$, $MC$, and $EMC$ curves, and label both the market equilibrium ($Q_M$ and $P_M$) and the socially optimal equilibrium ($Q^*$ and $P^*$).

## Question 5
Graphically indicate the deadweight loss (DWL) caused by the negative externality on the graph.


## Question 6
Calculate the optimal Pigouvian tax per unit that would lead producers to produce the socially optimal quantity.



::: {.callout-tip collapse="true"}
### Show answer

**Algebra.**

**Market equilibrium (ignoring externality):** set $MB(Q)=MC(Q)$:  
  $70 - 2Q = 10 + 2Q \Rightarrow 60 = 4Q \Rightarrow \boxed{Q_M = 15}$.  
  Price: $\boxed{P_M = MB(15) = MC(15) = 40}$.

**Socially optimal equilibrium:** set $MB(Q)=SMC(Q)$:  
  $70 - 2Q = 10 + 4Q \Rightarrow 60 = 6Q \Rightarrow \boxed{Q^* = 10}$.  
  Consumer price at the social optimum (what demand is willing to pay): $\boxed{P^* = MB(10) = 50}$.

**Pigouvian tax:** set $t^{*} = EMC(Q^{*})$:  
  $t^{*} = EMC(Q^{*}) = 2Q^{*}. \Rightarrow t^{*} = 2Q^{*} \Rightarrow \boxed{t^{*} = 20}$.
  
```{r}
#| fig-align: center
#| fig-width: 8
#| fig-height: 8

# Functions
MB  <- function(Q) 70 - 2*Q
MCf <- function(Q) 10 + 2*Q
EMC <- function(Q) 2*Q
SMC <- function(Q) 10 + 4*Q

# Key quantities
Q_M <- 15; P_M <- MB(Q_M)
Q_s <- 10; P_s <- MB(Q_s)
tax_star <- EMC(Q_s)
P_prod_star <- MCf(Q_s)

# Define the demand, marginal cost, and social marginal cost functions
x <- seq(0, 80, 10)

ggplot(data = data.frame(x =x,
                         y = x)) +
  geom_point(aes(x=x,y=y), size = 0, color = 'white') +
  geom_hline(yintercept = 0) +
  geom_vline(xintercept = 0) +
  geom_line(data = data.frame(Q = seq(0,35,5), 
                              Demand = 70 - 2*seq(0,35,5)), 
            aes(x = Q, y = Demand, color = 'MB'),
            linewidth = 1.5) +
  geom_line(data = data.frame(Q = x, 
                              MC = 10 + 2*x), 
            aes(x = Q, y = MC, color = 'MC'),
            linewidth = 1.5) +
  geom_line(
    data = data.frame(Q = seq(0, 35, 5),
                      EMC = 2 * seq(0, 35, 5)),
    aes(x = Q, y = EMC, color = "EMC"),
            linewidth = 1.5
  ) +
  geom_line(data = data.frame(Q = seq(0, 25, 5), 
                              SMC = 10 + 4*seq(0,25,5)), 
            aes(x = Q, y = SMC, color = 'SMC'),
            linewidth = 1.5) +
  geom_point(data = data.frame(x = 15, y = 40), aes(x = x, y = y), size = 2) +
  geom_point(data = data.frame(x = 10, y = 50), aes(x = x, y = y), size = 2) +
  geom_text(data = data.frame(x = 15+2.5, y = 40), 
            aes(x = x, y = y),
            label = expression(E_M)) +
  geom_text(data = data.frame(x = 10+2.5, y = 50), 
            aes(x = x, y = y),
            label = expression(E^"*")) +
  geom_polygon(data = data.frame(q = c(10, 15, 15), 
                                 p = c(50, 40, 70)),
               aes(x = q, y = p), 
               fill = "grey", alpha = 0.5) +
  geom_text(data = data.frame(x = 37.5, y = 50), 
            aes(x = x, y = y),
            label = "DWL from\n Overproduction") +
  geom_curve(aes(x = 14, y = 56, xend = 30, yend = 50), 
             curvature = 0.3, arrow = arrow(length = unit(0.03, "npc")),
             size = .5) +
  
  geom_segment(aes(x = Q_s, xend = Q_s, y = P_prod_star, yend = P_s),
               linewidth = 1) +
  geom_text(aes(x = Q_s - 1, y = (P_prod_star + P_s)/2),
            label = "t* = 20", angle = 90, vjust = 0.5, hjust = 1.1) +
  scale_x_continuous(breaks = seq(0,80,10)) +
  scale_y_continuous(breaks = seq(0,80,10)) +
  scale_color_manual("", 
                     breaks = c("MB", "MC", "EMC", "SMC"),
                     values = c("MB" = "blue", 
                                "MC" = "red", 
                                "EMC" = "pink", 
                                "SMC" = "darkred")) +
  labs(x = 'Q', y= 'P') +
  ylim(0,80) +
  guides(fill = guide_legend(title.position = "top",
                             label.position = "bottom",
                             keywidth = 3,
                             nrow = 1)) +
  theme(legend.position = c(.75, .5),
        axis.text.y = element_text(angle=90))
```

:::
