---
title: Classwork 4
subtitle: Common Property Resources
date: 2025-09-26
code-fold: true
execute: 
  echo: false
  eval: true
  warning: false
  message: false
from: markdown+emoji
---
```{r}
#| include: false

library(knitr)
library(rmarkdown)
library(tidyverse)
library(skimr)
library(ggthemes)
library(hrbrthemes)

theme_set(theme_ipsum() +
          theme(strip.background =element_rect(fill="lightgray"),
                axis.title.x = element_text(size = rel(1.5)),
                axis.title.y = element_text(size = rel(1.5)),
                legend.title = element_text(size=rel(1.25))
                ))
```


# Question 1 — Optimal Harvesting in a Fishery

Consider a fishery where the total benefit (**TB**) and total cost (**TC**) of fishing effort $E$ are:

- $TB(E)=1.5E - E^{2}$ (in \$)
- $TC(E)=0.5E$ (in \$)

where $E$ is the level of fishing effort (e.g., number of boat-days).

**a.** Draw the **Total Benefit (TB)** and **Total Cost (TC)** curves as functions of fishing effort $E$.  Place these curves in the **top panel** of a two-panel figure.


**b.** On the **bottom panel**, draw the **Marginal Benefit (MB)** and **Marginal Cost (MC)** curves as functions of fishing effort $E$.  

Make sure the top and bottom panels share the same horizontal scale for effort $E$, so the relationship between total and marginal spaces is consistent.

- Recall that:
  - $MB(E)=\dfrac{d\,TB}{dE}=1.5 - 2E$
  - $MC(E)=\dfrac{d\,TC}{dE}=0.5$


**c.** On both panels, **mark and label** the two benchmark effort levels using **vertical guide lines** that align across panels:

- The (economically) **efficient effort** $E^*$:
  Draw a vertical line at $E^*$ on **both** the top (TB/TC) and bottom (AB/AC/MB/MC) panels and label it “$E^*$”.

- The **open-access effort** $E_{OA}$:
  Draw a vertical line at $E_{OA}$ on **both** panels and label it “$E_{OA}$”.


```{r}
#| eval: false
# Load packages
library(tidyverse)

# Domain for effort (adjust as needed)
E_max <- 1.5
E <- seq(0, E_max, length.out = 501)

# Define functions
TB <- 1.5*E - E^2
TC <- 0.5*E
MB <- 1.5 - 2*E
MC_AC <- rep(0.5, length(E))
AB <- 1.5 - E


# Data frames
df_level <- tibble(E, TB, TC) |> pivot_longer(cols = c(TB, TC), names_to = "curve", values_to = "value")
df_marg  <- tibble(E, AB, MB, MC_AC) |> pivot_longer(cols = c(AB, MB, MC_AC), names_to = "curve", values_to = "value") |> mutate(curve = ifelse(curve == 'MC_AC', 'AC & MC', curve))

# (a) TB vs TC
ggplot(df_level, aes(x = E, y = value, linetype = curve)) +
  geom_line(size = 1) +
  labs(title = "Total Benefit and Total Cost vs Effort",
       x = "Effort (E)", y = "Dollars", linetype = "Curve") +
  theme_minimal(base_size = 12)

# (b) MB vs MC
ggplot(df_marg, aes(x = E, y = value, linetype = curve)) +
  geom_line(size = 1) +
  labs(title = "Average/Marginal Benefit and Average/Marginal Cost vs Effort",
       x = "Effort (E)", y = "Dollars per unit of E", linetype = "Curve") +
  theme_minimal(base_size = 12)
```
