---
title: Homework 2
subtitle: Externalities; Common Pool Resources; Public Goods
date: 2025-09-29
from: markdown+emoji
comments: false
echo: false
code-fold: false

toc: true
toc-depth: 6
toc-expand: 3

execute: 
  warning: false
  message: false
  fig-width: 9
  fig-height: 7
  fig-align: center

---
```{r}
#| include: false

library(tidyverse)
library(skimr)
library(DT)
library(hrbrthemes)
library(ggthemes)

theme_set(theme_ipsum() +
          theme(strip.background =element_rect(fill="lightgray"),
                axis.title.x = element_text(size = rel(1.5),),
                axis.title.y = element_text(size = rel(1.5),
                                            angle = 360),
                legend.title = element_text(size=rel(1.25))
                ))


df <- data.frame(
  q = seq(0,4,.25),
  mb_p = 4 - seq(0,4,.25),
  mc = seq(0,4,.25) + 2
) |> 
  pivot_longer(cols = mb_p:mc,
               names_to = 'type',
               values_to = 'p') |> 
  mutate(type = factor(type, labels = c('mb_p',
                                        'mc')),
         label = recode(type, 'mb_p' = "MB today", 'mc' = "MC"),
         group = interaction(type, label)  # Creating a unique identifier for color and linetype
  )

```



## Homework Instructions

- **No Generative AI**: You are **not** allowed to use generative AI tools for Homework Assignment 2, except for part **d2 in Question 1**

- **Deadlines**: **Wednesday, October 8, 2025 at 10:30 A.M.**  
    - Hand in your written answers on paper to Prof. Choe at the beginning of class.


# Question 1. Optimal Logging, Externalities, and REDD+ (Points: 50)

A forest jurisdiction evaluates logging activity $L$ under two private market components and one external cost:

- **Timber benefit (private):** $TB(L)=2L - L^{2}$  
- **Harvesting cost (private):** $TC(L)=L$  
- **Carbon damage (external):** $TD(L)=\delta\,L$, where $\delta = 0.5>0$ is the monetized climate damage per unit of logging effort $L$.

Here, $L$ is logging effort (e.g., worker-days). Private loggers ignore $TD$.

**Notes:**  
- $MB(L)=\frac{dTB}{dL}=2-2L$  
- $MC(L)=\frac{dTC}{dL}=1$  
- $AB(L)=TB/L=2-L$ (for $L>0$)  
- $AC(L)=TC/L=1$ (for $L>0$)  
- $MD(L)=\frac{dTD}{dL}=\delta=0.5$  


## a — Total Curves (Top Panel)
Plot **TB**, **TC**, **TD**, and **Total Social Cost** $TSC(L)=TC(L)+TD(L)$ versus $L$.


## b — Marginal Curves (Bottom Panel)
Plot **marginals**:  
- $MB(L)=2-2L$  
- $MC(L)=1$  
- $MD(L)=\delta=0.5$ (External Marginal Cost)

Also plot **averages**:  
- $AB(L)=2-L$  
- $AC(L)=1$


## c — Benchmarks (align vertical guide lines across panels)

For each benchmark below, (i) **denote private optimum $L^{priv}$, socially efficient optimum $L^{*}$, and open-access level $L_{OA}$**, (ii) **write the condition** to solve each, and (iii) **compute the numeric value for each** using $\delta=0.5$.  
Then **draw vertical guide lines** at the three effort levels on **both** panels so they align.

1. **Private optimum $L^{priv}$** — *Privately optimal with no externality considered.*  
   - *Definition:* A **single logging firm** (a private decision-maker with exclusive rights to the forest) maximizing their own net private benefit, **ignoring the external damage**.  

2. **Socially efficient optimum $L^{*}$** — *Maximizes total net social benefit (accounts for the external damage).*  
   - **Definition:** Chosen to internalize the carbon damage at the margin.  

3. **Open-access level $L_{OA}$** — *Free entry expands effort until profit is zero, ignoring externalities.*  
   - *Definition:* With **many competing loggers**, free entry expands effort until each logger earns zero economic profit and all resource rents are dissipated.


::: {.callout-tip collapse="true"}
### Show answer

```{r}
#| label: data
#| echo: false

delta <- 0.5                         # <-- adjust to explore
L_priv <- .5                    # MB = MC
L_star <- .25          # MB = MC + MD
L_OA   <- 1                    # AB = AC (ignoring TD)

L_max  <- 1.25
L      <- seq(0, L_max, length.out = 601)

TB  <- 2*L - L^2
TC  <- L
TD  <- delta*L
TSC <- TC + TD

MB <- 2 - 2*L
MC <- 1
MD <- rep(delta, length(L))

AB <- ifelse(L==0, NA_real_, TB/L)      # 2 - L
AC <- ifelse(L==0, NA_real_, TC/L)      # 0.2 + 0.3L

df_top <- tibble(L, TB, TC, TD, TSC) |>
  pivot_longer(-L, names_to = "curve", values_to = "value")

df_bot <- tibble(L, MB, MC, MD, AB) |>
  pivot_longer(-L, names_to = "curve", values_to = "value") |> 
  mutate(curve = ifelse(curve == "MC", "MC and AC", curve))
```

```{r}
#| label: top-panel
#| echo: false
#| fig-cap: "Top: Totals (TB, TC, TD, and TSC) vs logging effort L."
ggplot(df_top, aes(L, value, color = curve)) +
  geom_line(linewidth = 1) +
  geom_vline(xintercept = L_priv, linewidth = 0.6, alpha = 0.6, linetype = 2) +
  geom_vline(xintercept = L_star, linewidth = 0.6, alpha = 0.6, linetype = 2) +
  geom_vline(xintercept = L_OA,   linewidth = 0.6, alpha = 0.6, linetype = 2) +
  annotate("text", x = L_star, y = max(df_top$value)*0.95, label = "L*", vjust = -0.3) +
  annotate("text", x = L_priv, y = max(df_top$value)*0.95, label = "L_priv", vjust = -0.3) +
  annotate("text", x = L_OA,   y = max(df_top$value)*0.95, label = "L_OA", vjust = -0.3) +
  labs(title = "Totals: TB, TC, TD, and TSC vs Effort",
       x = "Logging Effort (L)", y = "Dollars", linetype = "Curve") +
  coord_cartesian(xlim = c(0, L_max)) +
  scale_x_continuous(breaks = seq(0,1.25,.25)) + 
  scale_color_tableau() +
  theme_minimal(base_size = 12)
```

```{r}
#| label: bottom-panel
#| fig-cap: "Bottom: Marginal & Average Curves (MB, MC, MD, AB, AC) vs L."
ggplot(df_bot, aes(L, value, color = curve)) +
  geom_line(linewidth = 1) +
  geom_vline(xintercept = L_priv, linewidth = 0.6, alpha = 0.6, linetype = 2) +
  geom_vline(xintercept = L_star, linewidth = 0.6, alpha = 0.6, linetype = 2) +
  geom_vline(xintercept = L_OA,   linewidth = 0.6, alpha = 0.6, linetype = 2) +
  annotate("text", x = L_star, y = max(df_bot$value, na.rm = TRUE)*0.95, label = "L*", vjust = -0.3) +
  annotate("text", x = L_priv, y = max(df_bot$value, na.rm = TRUE)*0.95, label = "L_priv", vjust = -0.3) +
  annotate("text", x = L_OA,   y = max(df_bot$value, na.rm = TRUE)*0.95, label = "L_OA", vjust = -0.3) +
  labs(title = "Marginals & Averages: MB, MC, MD, AB, AC",
       x = "Logging Effort (L)", y = "Dollars per unit of L", linetype = "Curve") +
  coord_cartesian(xlim = c(0, L_max),
                  ylim = c(0, 2)) +
  scale_x_continuous(breaks = seq(0,1.25,.25)) + 
  scale_color_tableau() +
  theme_minimal(base_size = 12)
```


:::


## d — What is REDD+ and how can it help here?

**REDD+** (*Reducing Emissions from Deforestation and forest Degradation*) is an international climate policy framework that provides payments when countries or jurisdictions **reduce logging and therefore avoid CO$_2$ emissions**.  

For simplicity, assume the climate damage from one unit of logging is valued at $\delta = 0.5$ dollars. Under REDD+, loggers (or local governments) receive a payment $\tau$ for each unit of logging **not carried out**, relative to a baseline.  

From the logger’s perspective, this acts like an **extra cost of logging**: cutting down one more unit of forest means losing $\tau$ dollars of REDD+ payment. In economic terms, the decision rule becomes  

$$
MB(L) = MC(L) + \tau.
$$

If $\tau$ is set equal to the external damage per unit of logging ($\delta$), then the logging decision takes into account the climate cost, and effort is reduced to the efficient level $L^*$. 

### d1 — Pigouvian Alignment
Assume $\delta=0.5$. Show that if a single logging firm faces a REDD+ payment set at $\tau=\delta$, it will choose the socially efficient effort $L^*$.



::: {.callout-tip collapse="true"}
### Show answer

- Setting $\tau=\delta$ makes the private firm's optimal condition a  $MB=MC+\delta$, which is equivalent to the condition for the socially efficient optimum. Therefore, setting $\tau=\delta$ implements $L^*$. 

:::


### d2 — Practical Challenges of REDD+
Briefly discuss the following implementation issues:

-	**Additionality**: How can we verify that emission reductions are truly extra (would not have happened without REDD+)?
-	**Leakage**: Could logging or deforestation simply shift to another location, offsetting the gains?
-	**Permanence**: Even if forests are protected now, what ensures the carbon remains stored long term (e.g., risk of future logging or fires)?



::: {.callout-tip collapse="true"}
### Show answer

REDD+ faces several implementation challenges that affect its credibility and long-term impact.  

- **Additionality:** REDD+ credits must reflect *real* emission reductions beyond what would have occurred without the program. Establishing a credible baseline ($L^{BAU}$) is difficult because future deforestation trends depend on market, policy, and social factors.  
- **Leakage:** Deforestation or logging reduced in one project area may simply shift to another unprotected area, offsetting the climate benefits. This risk is lower in jurisdictional or national-scale programs that track land-use changes more comprehensively.  
- **Permanence:** Even if forests are protected today, stored carbon can be released later through fires, illegal logging, or policy changes. Ensuring permanence requires long-term monitoring, insurance buffers, and legal safeguards to maintain carbon storage over time.  
Overall, addressing these three issues is essential to ensure REDD+ delivers genuine, measurable, and lasting climate benefits.

:::



<br><br>

# Question 2. Expansion of Data Centers and the Environment (Points: 50)


<div style="text-align: center; margin: 1em 0;">
  <iframe width="800" height="450" src="https://www.youtube.com/embed/t-8TDOFqkQA?si=oBBubL6CvNtGudUT" 
    title="YouTube video player" frameborder="0" allowfullscreen>
  </iframe>
  <p><em>Business Insider - Exposing The Dark Side of America's AI Data Center Explosion, September 18, 2025</em></p>
</div>

Watch the video above. Then solve the question:

## Video Summary

- **Rapid growth**: Data centers are being built at a rate of more than two per week, clustering near reliable power, water, cheap land, and tax breaks.
- **Scale & resources**: A single large center can use as much electricity and water as an entire city; clusters in Virginia, Arizona, and California consume enormous resources.
- **Externalities**:
  - *Local*: Noise, vibrations, health effects, and property value declines for nearby residents.
  - *Global*: High electricity demand driving fossil fuel use, CO₂ emissions, and massive water consumption in drought-stricken areas.
- **Opacity**: Companies frequently obscure ownership and operations by using shell LLCs (Limited Liability Companies), binding local officials with NDAs (Non-Disclosure Agreements), and claiming “trade secret” exemptions. Journalists had to reconstruct a public map of data centers by piecing together generator permits and scattered public records.
- **Equity & policy issues**:
  - Tax breaks and land-use approvals often benefit large technology companies, while the resulting costs (grid strain, water scarcity, higher household bills) are shifted onto local residents.
  - Even as companies promise “water positive” and “carbon neutral” goals, reliance on offsets and continued growth raises doubts.
  - Utilities in places like Virginia project billions in grid upgrades, with household electricity bills potentially rising 50%.
- **Employment mismatch**: Despite billions in investment, individual data centers create relatively few permanent jobs (sometimes fewer than 150).



## Grid Reliability as a Public Good

Two representative households, A and B, both benefit from **grid reliability** $G$.  
Grid reliability refers to the stability and resilience of the electricity system—fewer blackouts, quicker recovery from outages, and greater assurance that power is available when needed.  
This reliability is a **public good**: once provided, it benefits everyone simultaneously, and no household can be excluded from enjoying it.  

Each household’s **marginal benefit (MB)** function is:  

- $MB_A(G) = 40 - 2G$  
- $MB_B(G) = 30 - G$  

The **marginal cost (MC)** of providing additional reliability, such as building more transmission lines, adding backup generators, or investing in grid hardening, is given by:  

- $MC(G) = 10 + G$  

---

## a — Efficient Reliability Level
Find the efficient level $G^{*}$.

::: {.callout-tip collapse="true"}
### Show answer

**Samuelson condition:** Sum of marginal benefits equals marginal cost.
$$
MB_A(G) + MB_B(G) = MC(G)
$$

$$
(40 - 2G) + (30 - G) = 10 + G \;\Rightarrow\; 70 - 3G = 10 + G
$$

$$
60 = 4G \;\Rightarrow\; \boxed{G^* = 15}.
$$

*Interpretation:* At $G^*=15$, the additional social value of one more unit of reliability (adding $MB_A+MB_B$) exactly equals the marginal cost to provide it.
:::


---

## b — Private vs. Social Choice
Compare this to what might happen if households acted alone and only equated their **own $MB$** to $MC$.

::: {.callout-tip collapse="true"}
### Show answer

Because reliability is non-excludable, if a single household were the **sole** decision-maker, they would set **their own** $MB_i(G)$ equal to $MC(G)$, ignoring the other’s benefit:

- Household A alone:  
  $$
  40 - 2G = 10 + G \;\Rightarrow\; 30 = 3G \;\Rightarrow\; G_A = 10.
  $$

- Household B alone:  
  $$
  30 - G = 10 + G \;\Rightarrow\; 20 = 2G \;\Rightarrow\; G_B = 10.
  $$

Either household acting alone would choose $G=10$, which is **below** the efficient $G^*=15$, because each ignores the other’s marginal benefit. In voluntary settings with cost sharing, this logic produces **underprovision**: no one internalizes the full social gain from an extra unit of reliability.

*Key contrast:*  

- Socially efficient: uses **$MB_A + MB_B$** vs. $MC$ $\Rightarrow G^*=15$.  
- Private acting alone: uses **$MB_i$** vs. $MC$ $\Rightarrow G=10$.
:::


---

## c — Underprovision of Public Goods
Explain why **underprovision** arises.

::: {.callout-tip collapse="true"}
### Show answer

Because reliability is **non-excludable**, each household enjoys the benefits of any extra $G$ regardless of who pays. When deciding whether to pay for another unit, a private household compares **its own** $MB_i$ with $MC$ and ignores the other household’s benefit. As a result, the privately chosen level is where $MB_i=MC$, not where $MB_A+MB_B=MC$. This **free-rider problem** drives $G$ **below** the social optimum. In our numbers: private $G=10$ vs. efficient $G^*=15$.
:::


---

## d — Future Reliability with Data-Center Growth
Suppose data center growth increases society’s valuation of grid reliability and alters the cost of providing it. In the future:

- $MB_{A}^{Future}(G) = 60 - 2G$  
- $MB_{B}^{Future}(G) = 50 - G$  
- $MC^{Future}(G) = 10 + 2G$  

### d1 — Efficient Future Level 
Compute the efficient future level $G^{*Future}$.

::: {.callout-tip collapse="true"}
### Show answer

Apply Samuelson in the future scenario:
$$
(60 - 2G) + (50 - G) = 10 + 2G
$$
$$
110 - 3G = 10 + 2G \;\Rightarrow\; 100 = 5G \;\Rightarrow\; \boxed{G^{*Future} = 20}.
$$

*Comparison:* $G^{*}$ rises from $15$ to $20$ despite higher $MC$ slope, because the **increase in total marginal benefits** (from households valuing reliability more in a data-center-dependent economy) dominates.
:::


### d2 — Interpretation
Briefly interpret what the change in the optimal reliability level implies for society, considering:

- **Investment needs**: what new infrastructure or grid upgrades might be required?
-	**Costs**: how do higher marginal costs affect who pays (households vs. data centers)?
-	**Benefits**: how does increased reliability reduce risks (e.g., outages, health costs, lost data) for both residents and businesses?


::: {.callout-tip collapse="true"}
### Show answer

- **Investment needs:** A higher efficient level ($G^{*Future}=20$) justifies *more* grid investment: added transmission capacity, substation upgrades, storage/batteries, advanced protection systems, hardening against weather, and faster restoration capability.
- **Costs:** The future $MC$ is steeper ($10+2G$), so each additional unit of reliability is **more expensive**. Cost recovery design matters:
  - Use **non-distortionary components** (access/connection/capacity charges) for fixed/reliability costs, keep per-kWh prices near marginal energy cost.
  - Apply **cost-causation**: large new loads (e.g., data centers) should shoulder a proportionally larger share.
  - Pair with **equity protections** (discounted electricity prices or income-based credits) for vulnerable households.
- **Benefits:** Society is more **reliability-sensitive** (digital services, health, commerce). Higher $G$ reduces outage probability and severity, lowering losses from spoiled goods, lost data/transactions, health risks, and downtime for businesses and public services.

*Bottom line:* Even with higher marginal cost, the *social gains* from avoiding outages now justify a **higher** optimal reliability level.
:::


---

## f — Cost Sharing Debate
Who should bear the costs of grid expansion — households, data centers, or government? Justify.

::: {.callout-tip collapse="true"}
### Show answer

- **Principle: cost causation.** Entities that **drive** the need for expansion should cover commensurate shares (e.g., data centers with large peak demand and strict uptime requirements).
- **Households:** Benefit from reliability; contribute via broad, preferably **non-distortionary** components (e.g., modest fixed charges), with **means-tested protections**.
- **Data centers:** Large new loads should pay **higher interconnection and capacity charges**, reflecting their impact on peak capacity, redundancy, and transmission upgrades.
- **Government:** May support **public-good components** (broad resilience, cyber/black-start) via general funds if justified by economy-wide benefits.

**Recommended mix:** Apply **cost-causation charges** to large energy users, include **protections for low-income households**, and maintain **minimal distortion in per-kWh (volumetric) prices**. Address **externalities** (e.g., carbon pricing) **separately** from the recovery of non-energy fixed costs.

:::


---

## g — Local Government Incentives
Why might local governments still provide subsidies to large technology companies, or approve land-use changes that allow data centers to be built, even when these projects create negative externalities?

::: {.callout-tip collapse="true"}
### Show answer

- **Economic development narrative:** Large capital investment, construction jobs, and an expanded tax base are politically salient.
- **Prestige and competition:** Attracting big-name firms signals a “tech hub,” fueling a **race-to-the-bottom** across jurisdictions.
- **Fiscal optics:** Upfront incentives are visible wins; diffuse costs (higher bills, noise, water stress) are spread over residents and time.
- **Information asymmetries:** Non-disclosure agreements (NDAs), opaque ownership (LLCs), and “trade secret” claims can limit scrutiny of true costs.
- **Political cycles:** Short-term gains align with election timelines; externalities materialize later.

*Net effect:* Even when overall social welfare gains are limited, subsidies and approvals continue due to political incentives and regional competition for corporate projects.

:::


<br><br>

# Question 3. Class Participation (Not Graded for Homework 2)

Please write one sentence for each time you participated in class (any public speaking during class time) between **September 24, 2025 and October 6, 2025**.  

**Examples:**  

- “I asked a question to clarify the common-pool resource”  
- “I asked how market instruments affect consumer behavior.”  
- “I gave plastic pollution as an example of a negative externality.”  
- “I pointed out a calculation error during the lecture.”  
 


