---
title: Lecture 1
subtitle: Syllabus and Course Outline
format:
  clean-revealjs:
    self-contained: false
    chalkboard: true
    incremental: true
    code-annotations: hover
    scrollable: false
    css: _extensions/grantmcdermott/clean/clean.scss

    # logo: logo-title-slide.png
author:
  - name: Byeong-Hak Choe
    email: bchoe@geneseo.edu
    affiliations: SUNY Geneseo
date: 2025-08-25
execute: 
  eval: true
  echo: false
callout-icon: false

from: markdown+emoji
include-after-body: target-hover.html # effect.html

# bibliography: refs.bib
---

```{r setup}
#| include: false

library(tidyverse)
library(lubridate)
library(skimr)
library(ggthemes)
library(hrbrthemes)
library(viridis)
library(rmarkdown)
library(gapminder)
library(ggrepel)

theme_set(theme_fivethirtyeight() +
            theme(strip.background =element_rect(fill="lightgray"),
                axis.title.x = 
                  element_text(angle = 0,
                               size = rel(1.75),
                               margin = margin(10,0,0,0)),
                axis.title.y = 
                  element_text(angle = 0,
                               size = rel(1.75),
                               margin = margin(0,10,0,0)),
                axis.text.x = element_text(size = rel(1.75)),
                axis.text.y = element_text(size = rel(1.75)),
                strip.text = element_text(size = rel(1.5)),
                legend.position = "top",
                legend.text = element_text(size = rel(1.5)),
                legend.title = element_text(size = rel(1.5))
                )
          )

# Set global options for color-blind-friendly scales
# scale_colour_discrete <- function(...) scale_colour_viridis_d(...)
scale_colour_discrete <- function(...) scale_color_colorblind(...)
scale_fill_discrete <- function(...) scale_fill_colorblind(...)
```


# Instructor {background-color="#1c4982"}

## Instructor
### Current Appointment & Education

- Name: Byeong-Hak Choe.
- Assistant Professor of Data Analytics and Economics, School of Business at SUNY Geneseo.


- Ph.D. in Economics from University of Wyoming.
- M.S. in Economics from Arizona State University.
- M.A. in Economics from SUNY Stony Brook.
- B.A. in Economics & B.S. in Applied Mathematics from Hanyang University at Ansan, South Korea.
  - Minor in Business Administration.
  - Concentration in Finance.
  


## Instructor
### Economics and Data Science

- Choe, B.H., 2021. "**_Social Media Campaigns, Lobbying and Legislation: Evidence from #climatechange and Energy Lobbies._**"

- **Question**: To what extent do social media campaigns compete with fossil fuel lobbying on climate change legislation?

- Data include:
  - 5.0 million tweets with #climatechange/#globalwarming around the globe;
  - 12.0 million retweets/likes to those tweets;
  - 0.8 million Twitter users who wrote those tweets;
  - 1.4 million Twitter users who retweeted or liked those tweets;
  - 0.3 million US Twitter users with their location at a city level;
  - Firm-level lobbying data (expenses, targeted bills, etc.). 



## Instructor
### Economics and Data Science

- Choe, B.H. and Ore-Monago, T., 2024. "**_Governance and Climate Finance in the Developing World_**"

- **Climate finance** refers to the financial resources allocated for mitigating and adapting to climate change, including support for initiatives that reduce greenhouse gas emissions and enhance resilience to climate impacts.
  - We focus on *transnational financing* that rich countries provide poor countries with financial resources, in order to help them adapt to climate change and mitigate greenhouse gas (GHG) emissions.
  - Since the GHG emissions in developing countries are rapidly growing, it is crucial to assess the effectiveness of climate finance.
  - Poor governance (e.g., legal system, rule of law, and accountability) can be significant barriers to emissions reductions.


## Instructor
### Economics and Data Science

- Choe, B.H. and Newbold, Steve, "**_Estimating the Value of Statistical Life (VSL) through Big Data_**"

- VSL is the monetary value associated with reducing the risk of death. 
  - How much value would that be? How can we measure it?
  - How do government agencies use the VSL to decide which policies are worth the cost when they reduce the risk of death?





# Syllabus {background-color="#1c4982"}

## Syllabus

### Email, Class & Office Hours

- Email: [bchoe@geneseo.edu](bchoe@geneseo.edu)


- Class Websites: 
  1 [GitHub Course Website](http://bcecon.github.io/){target="_blank"}
  2 [Brightspace Course Shell](https://mylearning.suny.edu/d2l/home/2152120){target="_blank"}
  
  
- Office: South Hall 227B
- Office Hours: 
  - Mondays and Wednesdays 3:30 P.M.--5:00 P.M.
  - By appointment via email.


## Syllabus
### Course Description

- This course will focus on the role of the environment in the theory and practice of economics. 

- Emphasis is given to the role of **market failure** in explaining the existence of pollution, to alternative public policies for **pollution control** and **environmental management**, and to problems associated with **energy and climate change policies**.

- It will make use of microeconomic analysis at the intermediate level and will incorporate real-world examples. 


## Syllabus
### Course Description

- The class will be divided into three parts. 
  - Part I will cover **the ways in which markets fail** to efficiently allocate resources in the presence of pollution along with **environmental policies used to correct those failures**. 
  - Part II will focus on **energy policy** and the basics of **climate-related science, technology, and adaptation policies**. 
  - Part III will focus on **measuring the benefit of environmental protection**. 
  
- The course allows one to have a smart, intelligent, and informed conversation about environmental, energy, and climate change policies.




## Syllabus

### Course Learning Outcomes
- Understand the central **concepts and theories of environmental economics**.
- Frame an **environmental and energy policy** question from an **economist’s perspective**.
- Acquire a toolkit of methods for **environmental and energy economics and public policies**.
- Understand the economic drivers and implications of climate change in order to be able to describe **the whole purpose of the climate policies**.




## Syllabus

### Course Requirements
- **Homework**: 5 assignments
- **Quizzes**: 2 in-class quizzes
- **Participation**: In-person and online participation
- **Team Project**: *Climate Change and the Individual project* with team presentation and proposal
- **Exams**: 2 midterms and 1 final
  
## Syllabus
### Team Project – Climate Change and the Individual  

- Project topics should focus on **climate change actions, advocacy, awareness, and outreach**. 
  - For additional ideas, consider the [problems and possibilities](https://www.getitright.earth/problems-possibilities).
- Each team will consist of **three students**. 
- Teams must develop a clear **research question** for their term paper. 
- The proposal may be **theoretical, empirical, or a combination of both**. 
  - For empirical work, the proposed dataset must be approved in advance by **Byeong-Hak Choe**.

- Your team is encouraged to participate in the campus-wide [Ideas that Matter Student Challenge](https://www.geneseo.edu/cil/ideas-matter-student-challenge) in Spring 2026.


## Syllabus
### Expectations for the Proposal
Your team’s proposal should critically examine a climate change issue by: 

1. **Explaining** the central dilemma facing society.  
2. **Identifying** the key questions that must be addressed to develop solutions, while acknowledging the trade-offs between objectives.  
3. **Presenting** appropriate evidence to analyze the competing interests involved.  
4. **Formulating** policy recommendations that logically follow from your analysis.  


## Syllabus
### Required Materials

All required books are open access with free PDF files:

- [*An Introduction to Energy Economics andz Policy*](https://www.cambridge.org/core/services/aop-cambridge-core/content/view/330248A5D38A65BF463C59A26E193CB7/9781009471817AR.pdf/An_Introduction_to_Energy_Economics_and_Policy.pdf?event-type=FTLA), by Massimo Filippini and Suchita Srinivasan (Henthforce, FS)

- [*Microeconomics and the Environment*](https://www.bu.edu/eci/files/2024/07/MicroAndtheEnvironment-2024.pdf), by Brian Roach, Erin Lennox and Anne-Marie Codur (Henceforth, RLC)
  
- [*Energy Economics and Policy*](https://www.bu.edu/eci/files/2025/06/Energy-Economics-and-Policy-Module-2025.pdf), by Brian Roach and Jonathan M. Harris (Henceforth, RH)

- [*The Economics of Renewable Energy*](https://www.bu.edu/eci/files/2024/03/renewable-energy-econ-module-final.pdf), by David Timmons, Jonathan Harris and Brian Roach (Henceforth, THR)

- [*The Economics of Global Climate Change*](https://www.bu.edu/eci/files/2025/07/Climate-Module-2025-final.pdf), by Jonathan M. Harris, Brian Roach and Anne-Marie Codur (Henceforth, HRC)


## Syllabus
### Recommended References

- [*Global Climate Change: The Pragmatist’s Guide to Moving the Needle*](https://climate-change.ivo-welch.info/home/), by Ivo Welch and Bradford Cornell  
- [*What If We Get It Right?: Visions of Climate Futures*](https://www.getitright.earth), by Ayana Elizabeth Johnson  
- [*An Introduction to Natural Disaster Economics*](https://docs.google.com/document/d/1mh6ltbYDUguee8JkPfPGGfM26Lz7cEa7/edit?usp=sharing&ouid=113346383464248167416&rtpof=true&sd=true), by Matthew Khan  
- **Markets and the Environment**, by Nathaniel O. Keohane and Sheila M. Olmstead (2nd Edition)   


## Syllabus
### Recommended References

- **_Economics and the Environment_** by Eban S. Goodstein, Jason Wong, and Stephen Polasky (10th Edition; Henthforce, GWP)  
- __*Environmental and Natural Resource Economics*__, by Tom Tietenberg and Lynne Lewis (12th Edition; Henthforce, TL)  
- __*Introduction to Environmental Economics*__, by Nick Hanley, Jason Shogren, and Ben White (3rd Edition; Henthforce, HSW)  
- __*Environmental and Natural Resource Economics: A Contemporary Approach*__, by Jonathan Harris and Brian Roach (5th Edition)  



## Syllabus
### Course Contents and Schedule

```{r, out.width='100%', fig.align='center'}
#| eval: true
#| echo: false
text_tbl <- data.frame(
  Week = c("1", "2-5", "6-7"),
  Content = c("Introduction; Microeconomics Review",
"Market Failures and Behavioral Anomalies",
"Public Goods and Common Pool Resources"),
  `Material` = c("FS Ch.1; RLC Ch.1-2; HRC", "FS Ch.2; RLC Ch2-3", "FS Ch.2; RLC Ch2-3")
  )


# Create a DT datatable without search box and 'Show entries' dropdown
DT::datatable(text_tbl, rownames = FALSE,
              options = list(
  dom = 't', # This sets the DOM layout without the search box ('f') and 'Show entries' dropdown ('l')
  paging = FALSE, # Disable pagination
  columnDefs = list(list(
    targets = "_all", # Applies to all columns
    orderable = FALSE # Disables sorting
  ))
), callback = htmlwidgets::JS("
  // Change header background and text color
  $('thead th').css('background-color', '#1c4982');
  $('thead th').css('color', 'white');

  // Loop through each row and alternate background color
  $('tbody tr').each(function(index) {
    if (index % 2 == 0) {
      $(this).css('background-color', '#d1dae6'); // Light color for even rows
    } else {
      $(this).css('background-color', '#9fb2cb'); // Dark color for odd rows
    }
  });

  // Set text color for all rows
  $('tbody tr').css('color', 'black');

  // Add hover effect
  $('tbody tr').hover(
    function() {
      $(this).css('background-color', '#607fa7'); // Color when mouse hovers over a row
    }, 
    function() {
      var index = $(this).index();
      if (index % 2 == 0) {
        $(this).css('background-color', '#d1dae6'); // Restore even row color
      } else {
        $(this).css('background-color', '#9fb2cb'); // Restore odd row color
      }
    }
  );
")
)

```




## Syllabus
### Course Contents and Schedule
```{r, out.width='100%', fig.align='center'}
#| eval: true
#| echo: false
text_tbl <- data.frame(
  Week = c("8", "9-11", "12", "13-14", "15"),
  Content = c(
"Economics of Global Climate Change",
"Energy Policy",
"Climate Change Adaptation: Floods and Wildfire",
"Measuring the Benefits of Environmental Protection",
"Team Presentations"
),
  `Material` = c("HRC", "FS Ch.7-8; RH; THR", "TL Ch.14", "GWP Ch.5; HSW Ch5", "")
  )

# Create a DT datatable without search box and 'Show entries' dropdown
DT::datatable(text_tbl, rownames = FALSE,
              options = list(
  dom = 't', # This sets the DOM layout without the search box ('f') and 'Show entries' dropdown ('l')
  paging = FALSE, # Disable pagination
  columnDefs = list(list(
    targets = "_all", # Applies to all columns
    orderable = FALSE # Disables sorting
  ))
), callback = htmlwidgets::JS("
  // Change header background and text color
  $('thead th').css('background-color', '#1c4982');
  $('thead th').css('color', 'white');

  // Loop through each row and alternate background color
  $('tbody tr').each(function(index) {
    if (index % 2 == 0) {
      $(this).css('background-color', '#d1dae6'); // Light color for even rows
    } else {
      $(this).css('background-color', '#9fb2cb'); // Dark color for odd rows
    }
  });

  // Set text color for all rows
  $('tbody tr').css('color', 'black');

  // Add hover effect
  $('tbody tr').hover(
    function() {
      $(this).css('background-color', '#607fa7'); // Color when mouse hovers over a row
    }, 
    function() {
      var index = $(this).index();
      if (index % 2 == 0) {
        $(this).css('background-color', '#d1dae6'); // Restore even row color
      } else {
        $(this).css('background-color', '#9fb2cb'); // Restore odd row color
      }
    }
  );
")
)

```



## Syllabus
### Class Schedule

- **No Classes**
  - Week 2 (Sep 1): **Labor Day**
  - Week 8 (Oct 13–14): **Fall Break**
  - Week 14 (Nov 26–28): **Thanksgiving Break**



## Syllabus
### Class Schedule

- **Homework, Quiz, Exam & Team Project**
  - Week 4 (Sep 15–19): **Homework 1**: 
  - Week 7 (Oct 6–10): **Homework 2, Quiz 1 & Midterm Exam 1**
  - Week 10 (October 31, Friday, 11:59 P.M.): **Project Kickoff Report**
  - Week 11 (Nov 3–7): **Homework 3**
  - Week 13 (Nov 17–21): **Homework 4, Quiz 2, & Midterm Exam 2**
  - Week 15 (Dec 1–5): **Team Presentations**
  - Week 16 (Dec 8-12): **Homework 5**
  - Week 16 (December 10, Wednesday, 8:30 A.M.-10:30 A.M.): **Final Exam**
  - Week 17 (December 16, Tuesday, 11:59 P.M.): **Project Proposal**



## Syllabus
### Grading

$$
\begin{align}
\quad\\
&\text{(Total Percentage Score)} \\
= &\;\;\;\, 0.05 \times \text{(Attendance)} \\
&+ 0.05 \times \text{(Quiz & Participation)}\\
&+ 0.20 \times \text{(Homework)}\\
&+ 0.20 \times \text{(Team Project)}\\
&+ 0.50 \times \text{(Exams)}
\end{align}
$$



## Syllabus
### Grading - Attendance & Homework

- You are allowed up to **6 absences** without penalty. 
  - Send me an email if you have standard excused reasons (illness, family emergency, transportation problems, etc.).

- For each absence beyond the initial six, there will be a deduction of 1% point from the Total Percentage Grade.

- The single lowest homework score will be dropped when calculating the total homework score. 
  - Each homework except for the homework with the lowest score accounts for 25% of the total homework score.


## Syllabus
### Grading

$$
\begin{align}
&\quad(\text{Midterm Exam Score})\\
&= \max\left\{0.50 \times \text{(Midterm 1 Score)} + 0.50 \times \text{(Midterm 2 Score)},\right.\\ 
&\qquad\quad\;\,\left.0.33 \times \text{(Midterm 1 Score)} + 0.67 \times \text{(Midterm 2 Score)}\right\}.
\end{align}
$$

- The Midterm Exam Score will be the **higher** of the following two calculations:

  1. The simple average of Midterm Exam 1 and Midterm Exam 2
  2. The weighted average, with one-third weight on Midterm Exam 1 and two-thirds weight on Midterm Exam 2

- Midterm Exam 2 is cumulative.

## Syllabus
### Grading

$$
\begin{align}
&(\text{Total Exam Score}) \\
=\, &\text{max}\,\left\{0.50\times(\text{Midterm Exam Score}) \,+\, 0.50\times(\text{Final Exam Score})\right.,\notag\\ 
&\qquad\;\,\left.0.25\times(\text{Midterm Exam Score}) \,+\, 0.75\times(\text{Final Exam Score})\right\}.\notag
\end{align}
$$


- The Total Exam Score will be the **higher** of the following two calculations:

  1. The simple average of Midterm Exam and Final Exam
  2. The weighted average, with one-fourth weight on Midterm Exam and three-forth weight on Final Exam.

- Final Exam is cumulative.

## Syllabus
### Make-up Policy

-  Make-up exams will not be given unless you have either a medically verified excuse or an absence excused by the University.

- If you cannot take exams because of religious obligations, notify me by email at least two weeks in advance so that an alternative exam time may be set.

- A missed exam without an excused absence earns a grade of zero.

- Late submissions for homework assignment will be accepted with a penalty. 

- A zero will be recorded for a missed assignment.


## Syllabus
### Academic Integrity and Plagiarism

-  All homework assignments and exams must be the original work by you. 

- Examples of academic dishonesty include:
  - *Representing the work, thoughts, and ideas of another person as your own*
  - *Allowing others to represent your work, thoughts, or ideas as theirs*, and
  - *Being complicit in academic dishonesty by suspecting or knowing of it and not taking action*.


- Geneseo’s Library offers frequent workshops to help you understand how to **paraphrase**, **quote**, and **cite** outside sources properly. 
  - See [https://www.geneseo.edu/library/library-workshops](https://www.geneseo.edu/library/library-workshops){target="_blank"}.


## Syllabus
### Artificial Intelligence (AI) Policy

- Unless AI tools are explicitly permitted for homework or in-class quizzes, you must complete your work independently. 

- This means you should not use tools like ChatGPT for any aspect of our coursework. 

- Such use is a form of academic dishonesty. Use of such tools is not only cheating, it will also cheat you of the opportunity to learn and develop your own skills. 

- While AI will undoubtedly play important roles in our future society, you will be better able to utilize AI if you have developed your own critical thinking, writing, and analytical skills by doing your own work. 

- If you have any questions about this, please ask.



## Syllabus
### Accessibility

-  The Office of Accessibility will coordinate reasonable accommodations for persons with physical, emotional, or cognitive disabilities to ensure equal access to academic programs, activities, and services at Geneseo.

- Please contact me and the Office of Accessibility Services for questions related to access and accommodations.



## Syllabus
### Career Design

- To get information about career development, you can visit the Career Development Events Calendar ([https://www.geneseo.edu/career_development/events/calendar](https://www.geneseo.edu/career_development/events/calendar){target="_blank"}).

- You can stop by South 112 to get assistance in completing your Handshake Profile [https://app.joinhandshake.com/login](https://app.joinhandshake.com/login){target="_blank"}.
  - Handshake is ranked #1 by students as the best place to find full-time jobs.
  - 50% of the 2018-2020 graduates received a job or internship offer on Handshake.
  - Handshake is trusted by all 500 of the Fortune 500.

