---
title: Lecture 1
subtitle: Syllabus, Course Outline, and Introduction
format:
  clean-revealjs:
    self-contained: false
    chalkboard: true
    incremental: true
    code-annotations: hover
    scrollable: false

    # logo: logo-title-slide.png
author:
  - name: Byeong-Hak Choe
    email: bchoe@geneseo.edu
    affiliations: SUNY Geneseo
date: 2024-08-26
execute: 
  eval: true
  echo: true
callout-icon: false

from: markdown+emoji
include-after-body: target-hover.html # effect.html

# bibliography: refs.bib
---


```{r setup}
#| include: false
library(tidyverse)
library(skimr)
library(ggthemes)
library(hrbrthemes)


theme_set(theme_fivethirtyeight()+
          theme(strip.background =element_rect(fill="lightgray"),
                axis.title.x = 
                  element_text(angle = 0,
                               size = rel(1.5),
                               margin = margin(10,0,0,0)),
                axis.title.y = 
                  element_text(angle = 0,
                               size = rel(1.5),
                               margin = margin(0,10,0,0)),
                axis.text.x = element_text(size = rel(1.5)),
                axis.text.y = element_text(size = rel(1.5)),
                strip.text = element_text(size = rel(1.5)),
                legend.position = "top",
                legend.text = element_text(size = rel(1.5)),
                legend.title = element_text(size = rel(1.5))
                )
          )
```


# Instructor {background-color="#1c4982"}

## Instructor
### Current Appointment & Education

- Name: Byeong-Hak Choe.
- Assistant Professor of Data Analytics and Economics, School of Business at SUNY Geneseo.


- Ph.D. in Economics from University of Wyoming.
- M.S. in Economics from Arizona State University.
- M.A. in Economics from SUNY Stony Brook.
- B.A. in Economics & B.S. in Applied Mathematics from Hanyang University at Ansan, South Korea.
  - Minor in Business Administration.
  - Concentration in Finance.
  


## Instructor
### Economics and Data Science

- Choe, B.H., 2021. "Social Media Campaigns, Lobbying and Legislation: Evidence from #climatechange and Energy Lobbies."

- Question: To what extent do social media campaigns compete with fossil fuel lobbying on climate change legislation?

- Data include:
  - 5.0 million tweets with #climatechange/#globalwarming around the globe;
  - 12.0 million retweets/likes to those tweets;
  - 0.8 million Twitter users who wrote those tweets;
  - 1.4 million Twitter users who retweeted or liked those tweets;
  - 0.3 million US Twitter users with their location at a city level;
  - Firm-level lobbying data (expenses, targeted bills, etc.). 



## Instructor
### Economics and Data Science

- Choe, B.H. and Ore-Monago, T., 2024. "Governance and Climate Finance in the Developing World"

- **Climate finance** refers to the financial resources allocated for mitigating and adapting to climate change, including support for initiatives that reduce greenhouse gas emissions and enhance resilience to climate impacts.
  - We focus on transnational financing that rich countries provide poor countries with financial resources, in order to help them adapt to climate change and mitigate greenhouse gas (GHG) emissions.
  - Since the GHG emissions in developing countries are rapidly growing, it is crucial to assess the effectiveness of climate finance.
  - Poor governance can be significant barriers to emissions reductions within these countries.


## Instructor
### Economics and Data Science

- Choe, B.H. and Ore-Monago, T., 2024. "Governance and Climate Finance in the Developing World"

- Data include:
  - Global climate finance data (e.g., donors, recipients, characteristics of climate change projects)
  - World Bank Governance Indicators over the years (e.g., government effectiveness, voice and accountability, political stability and absence of violence/terrorism, regulatory quality, rule of law, control of corruption)
  - Various economic indicators (e.g., trade pattern of low carbon technology products, macroeconomic risks, energy)


## Instructor
### Economics and Data Science

- Choe, B.H., James, Alex and Newbold, Steve, 2025. "Estimating the Value of Statistical Life (VSL) through Big Data"

- VSL is the monetary value associated with reducing the risk of death. 
  - How much value would that be? How can we measure it?




# Syllabus {background-color="#1c4982"}

## Syllabus

### Email, Class & Office Hours

- Email: [bchoe@geneseo.edu](bchoe@geneseo.edu)


- Class Homepage: 
  i. [https://mylearning.suny.edu/d2l/home/1513939](https://mylearning.suny.edu/d2l/home/1513939){target="_blank"}
  i. [http://bcecon.github.io/](http://bcecon.github.io/){target="_blank"}
  
  
- Office: South Hall 301
- Office Hours: 
  - In-person: Mondays and Wednesdays 1:30 P.M.--3:00 P.M.
  - Online: by appointment via email.


## Syllabus

### Course Description

- This course examines the role of economic analysis in addressing challenges in natural resource management. 
- We will begin by reviewing key economic concepts, for example, externalities and market failures, benefit-cost analysis, and non-market valuation techniques. 

- We will then apply them to issues concerning various natural resources, such as forests and fisheries. 

- Additionally, we will explore climate change topics, including energy policy, carbon pricing, and adaptation strategies. 

- A major focus is on assessing the optimal role of public policy. 




## Syllabus

### Course Learning Outcomes

- Grasp the fundamental concepts and theories of natural resource economics, including scarcity, allocation, externalities, sustainability, and ecosystem service and goods.
- Analyze natural resource management issues using economic principles to evaluate trade-offs and policy options.
- Understand and be able to use benefit-cost analysis.
- Understand the economic drivers and implications of climate change, enabling clear articulation of its significance in resource management and policy discussions.
- Understand the role of market and non-market mechanisms in natural resource allocation and management. 
- Explore the economic aspects of renewable and non-renewable resource management, including energy, forest, and fisheries.


## Syllabus

### Required Materials

- **Environmental and Natural Resource Economics** (12th Edition) by Tom Tietenberg and Lynne Lewis (Henthforth, TL)

- Additional readings will be provided.


## Syllabus

### Course Requirements
- **Homework**: There will be six homework assignments.
- **Quiz**: There will be in-class quizzes.
- **Class Participation**: You are encouraged to ask questions and discuss topics during the class time.
- **Exam**: There will be two short midterm exams and one comprehensive final exam.
- **Term paper**: There will be one term paper.
  
  
  
## Syllabus

### Term paper

- Your term paper should critically examine natural resource management issues by: 
  1. Thoroughly explaining the central dilemma facing society, 
  2. Identifying the key questions that must be addressed to develop solutions, while recognizing the trade-offs between different objectives, 
  3. Presenting appropriate evidence to analyze the competing interests involved, and 
  4. Formulating **policy recommendations** that logically follow from your analysis.

- A list of suggested topics for the term paper will be provided. However, each team may also choose a topic of its own that is related to natural resource management issues.


## Syllabus

### Term paper
- Each team for the term paper consists of one or two student(s). 
- Each team must develop a research question for the term paper. 
- The term paper can be either theoretical, empirical, or both. 
- For empirical research, the data for the term paper must be approved by Byeong-Hak Choe. 
- Term papers will be evaluated based on (1) the quality of research question, (2) the quality of oral presentation, (3) the quality of economic analysis, (4) the quality of modeling analysis, and (5) peer evaluation on other students' presentations. 
- Changing a team member or a research topic must be approved by Byeong-Hak Choe.



## Syllabus

### Class Schedule

- There will be tentatively 42 class sessions.
  - **The midterm exam 1** is scheduled on **October 9, 2024, Wednesday**, during the class time.
  - The due for the **term paper proposal** is **October 31, 2024, Thursday**, 11:59 P.M.
  - **The midterm exam 2** is scheduled on **November 8, 2024, Friday**, during the class time.
  - **The presentation for the term paper** is scheduled on **December 9, 2024, Monday**, during the class time.
  - **The final exam** is scheduled on **December 13, 2024, Friday, 3:30 P.M.--5:30 P.M.**
  - The due for **the term paper** is **December 17, 2024, Tuesday, 11:59 P.M.**




## Syllabus
### Course Contents

```{r, out.width='100%', fig.align='center'}
#| eval: true
#| echo: false
text_tbl <- data.frame(
  Week = c("1", "2", "3", "4", "5", "6", "7"),
  Contents = c("Course Outlines; Vision of the Future",
"The Economic Approach: Property Rights, Externalities, and Environmental Problems",
"Evaluating Trade-Offs: Benefit-Coast Analysis and Other Decision-Making Metrics",
"Valuing the Environment: Methods",
"Valuing the Environment: Methods",
"Dynamic Efficiency and Sustainable Development", "Water Pollution: Managing Water Quality for Rivers, Lakes, and Oceans"),
  `HW Exam` = c("", "", "HW 1", "", "", "HW 2", "Midterm I")
  )


# Create a DT datatable without search box and 'Show entries' dropdown
DT::datatable(text_tbl, rownames = FALSE,
              options = list(
  dom = 't', # This sets the DOM layout without the search box ('f') and 'Show entries' dropdown ('l')
  paging = FALSE, # Disable pagination
  columnDefs = list(list(
    targets = "_all", # Applies to all columns
    orderable = FALSE # Disables sorting
  ))
), callback = htmlwidgets::JS("
  // Change header background and text color
  $('thead th').css('background-color', '#1c4982');
  $('thead th').css('color', 'white');

  // Loop through each row and alternate background color
  $('tbody tr').each(function(index) {
    if (index % 2 == 0) {
      $(this).css('background-color', '#d1dae6'); // Light color for even rows
    } else {
      $(this).css('background-color', '#9fb2cb'); // Dark color for odd rows
    }
  });

  // Set text color for all rows
  $('tbody tr').css('color', 'black');

  // Add hover effect
  $('tbody tr').hover(
    function() {
      $(this).css('background-color', '#607fa7'); // Color when mouse hovers over a row
    }, 
    function() {
      var index = $(this).index();
      if (index % 2 == 0) {
        $(this).css('background-color', '#d1dae6'); // Restore even row color
      } else {
        $(this).css('background-color', '#9fb2cb'); // Restore odd row color
      }
    }
  );
")
)

```




## Syllabus
### Course Contents

```{r, out.width='100%', fig.align='center'}
#| eval: true
#| echo: false
text_tbl <- data.frame(
  Week = c("8", "9", "10", "11"),
  Contents = c(
"Climate Change I: The Nature of the Challenge",
"Climate Change II: The Role of Energy Policy",
"Climate Change III: Carbon Pricing",
"Climate Change IV: Adaptation: Floods, Wildfires, and Water Scarcity"
),
  `HW Exam` = c("", "HW 3", "", "HW 4 & Midterm II")
  )


# Create a DT datatable without search box and 'Show entries' dropdown
DT::datatable(text_tbl, rownames = FALSE,
              options = list(
  dom = 't', # This sets the DOM layout without the search box ('f') and 'Show entries' dropdown ('l')
  paging = FALSE, # Disable pagination
  columnDefs = list(list(
    targets = "_all", # Applies to all columns
    orderable = FALSE # Disables sorting
  ))
), callback = htmlwidgets::JS("
  // Change header background and text color
  $('thead th').css('background-color', '#1c4982');
  $('thead th').css('color', 'white');

  // Loop through each row and alternate background color
  $('tbody tr').each(function(index) {
    if (index % 2 == 0) {
      $(this).css('background-color', '#d1dae6'); // Light color for even rows
    } else {
      $(this).css('background-color', '#9fb2cb'); // Dark color for odd rows
    }
  });

  // Set text color for all rows
  $('tbody tr').css('color', 'black');

  // Add hover effect
  $('tbody tr').hover(
    function() {
      $(this).css('background-color', '#607fa7'); // Color when mouse hovers over a row
    }, 
    function() {
      var index = $(this).index();
      if (index % 2 == 0) {
        $(this).css('background-color', '#d1dae6'); // Restore even row color
      } else {
        $(this).css('background-color', '#9fb2cb'); // Restore odd row color
      }
    }
  );
")
)

```




## Syllabus
### Course Contents

```{r, out.width='100%', fig.align='center'}
#| eval: true
#| echo: false
text_tbl <- data.frame(
  Week = c("12", "13", "14", "15", "16"),
  Contents = c(
"Ecosystem Services: Nature’s Threatened Bounty",
"Common-Pool Resources: Commercially Valuable Fisheries",
"Forests: Storable, Renewable Resources",
"Land: A Locationally Fixed, Multipurpose Resource", "Term Paper Presentation"
),
  `HW Exam` = c("", "HW 5", "", "HW 6", "Final & Term Paper")
  )


# Create a DT datatable without search box and 'Show entries' dropdown
DT::datatable(text_tbl, rownames = FALSE,
              options = list(
  dom = 't', # This sets the DOM layout without the search box ('f') and 'Show entries' dropdown ('l')
  paging = FALSE, # Disable pagination
  columnDefs = list(list(
    targets = "_all", # Applies to all columns
    orderable = FALSE # Disables sorting
  ))
), callback = htmlwidgets::JS("
  // Change header background and text color
  $('thead th').css('background-color', '#1c4982');
  $('thead th').css('color', 'white');

  // Loop through each row and alternate background color
  $('tbody tr').each(function(index) {
    if (index % 2 == 0) {
      $(this).css('background-color', '#d1dae6'); // Light color for even rows
    } else {
      $(this).css('background-color', '#9fb2cb'); // Dark color for odd rows
    }
  });

  // Set text color for all rows
  $('tbody tr').css('color', 'black');

  // Add hover effect
  $('tbody tr').hover(
    function() {
      $(this).css('background-color', '#607fa7'); // Color when mouse hovers over a row
    }, 
    function() {
      var index = $(this).index();
      if (index % 2 == 0) {
        $(this).css('background-color', '#d1dae6'); // Restore even row color
      } else {
        $(this).css('background-color', '#9fb2cb'); // Restore odd row color
      }
    }
  );
")
)

```



## Syllabus
### Grading

$$
\begin{align}
(\text{Total Percentage Grade}) =&\;\, 0.05\times(\text{Total Attendance}) \notag\\
&\,+\, 0.15\times(\text{Total Quiz & Class Participation})\notag\\ 
& \,+\, 0.15\times(\text{Total Homework})\notag\\  
&\,+\, 0.15\times(\text{Term Paper})\notag\\
& \,+\, 0.50\times(\text{Total Exam}).\notag
\end{align}
$$



## Syllabus
### Grading



- You are allowed up to 6 absences without penalty. 
  - Send me an email if you have standard excused reasons (illness, family emergency, transportation problems, etc.).

- For each absence beyond the initial six, there will be a deduction of 1% from the Total Percentage Grade.

- The single lowest homework score will be dropped when calculating the total homework score. 
  - Each homework except for the homework with the lowest score accounts for 20% of the total homework score.


## Syllabus
### Grading


- The total exam score is the maximum between 
  1. the simple average of the midterm exam score and the final exam score and
  2. the weighted average of them with one-fourth weight on the midterm exam score and three-third weight on the final exam score:

$$
\begin{align}
&(\text{Total Exam Score}) \\
=\, &\text{max}\,\left\{0.50\times(\text{Midterm Exam Score}) \,+\, 0.50\times(\text{Final Exam Score})\right.,\notag\\ 
&\qquad\;\,\left.0.25\times(\text{Midterm Exam Score}) \,+\, 0.75\times(\text{Final Exam Score})\right\}.\notag
\end{align}
$$



## Syllabus
### Make-up Policy

-  Make-up exams will not be given unless you have either a medically verified excuse or an absence excused by the University.

- If you cannot take exams because of religious obligations, notify me by email at least two weeks in advance so that an alternative exam time may be set.

- A missed exam without an excused absence earns a grade of zero.

- Late submissions for homework assignment will be accepted with a penalty. 

- A zero will be recorded for a missed assignment.


## Syllabus
### Academic Integrity and Plagiarism

-  All homework assignments and exams must be the original work by you. 

- Examples of academic dishonesty include:
  - *representing the work, thoughts, and ideas of another person as your own*
  - *allowing others to represent your work, thoughts, or ideas as theirs*, and
  - *being complicit in academic dishonesty by suspecting or knowing of it and not taking action*.


- Geneseo’s Library offers frequent workshops to help you understand how to **paraphrase**, **quote**, and **cite** outside sources properly. 
  - See [https://www.geneseo.edu/library/library-workshops](https://www.geneseo.edu/library/library-workshops){target="_blank"}.


## Syllabus
### Artificial Intelligence (AI) Policy

- Unless AI tools are explicitly permitted for homework or in-class quizzes, you must complete your work independently. 

- This means you should not use tools like ChatGPT for any aspect of our coursework. 

- Such use is a form of academic dishonesty. Use of such tools is not only cheating, it will also cheat you of the opportunity to learn and develop your own skills. 

- While AI will undoubtedly play important roles in our future society, you will be better able to utilize AI if you have developed your own critical thinking, writing, and analytical skills by doing your own work. 

- If you have any questions about this, please ask.



## Syllabus
### Accessibility

-  The Office of Accessibility will coordinate reasonable accommodations for persons with physical, emotional, or cognitive disabilities to ensure equal access to academic programs, activities, and services at Geneseo.

- Please contact me and the Office of Accessibility Services for questions related to access and accommodations.



## Syllabus
### Career Design

- To get information about career development, you can visit the Career Development Events Calendar ([https://www.geneseo.edu/career_development/events/calendar](https://www.geneseo.edu/career_development/events/calendar){target="_blank"}).

- You can stop by South 112 to get assistance in completing your Handshake Profile [https://app.joinhandshake.com/login](https://app.joinhandshake.com/login){target="_blank"}.
  - Handshake is ranked #1 by students as the best place to find full-time jobs.
  - 50% of the 2018-2020 graduates received a job or internship offer on Handshake.
  - Handshake is trusted by all 500 of the Fortune 500.

